### 2011-02-11 Text/paper.tex
Copied results. Next step is to obtain steady solutions, non-diagonal dependency and capillary curve.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Recently we investigated the applicability and parameters choice for the two-dimensional case
Bretherton problem using the lattice Boltzmann method (LBM) {\color{red} Give citation here}. This
paper is the continuation of the previous work and is focused on the three-dimensional case to
validate the LBM for simulation of phenomena for rectangular and square capillaries in the
moderate range of Capillary number parameters. 
\end{abstract}
\section{What to examine}
I suggest to examine first of all the application of the LBM for really small capillary number,
where it's not possible to resolve the interface. In this case one should find the asymmetric
radiuses for the droplet, the difference should be around $10$ percent. While applying the forcing
it should come to the spheric shape. As well we need to examine the capillary number dependancy on
the length and suggest to examine some axicentric factors for rectangular channels.

\section{Introduction}
While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-bretherton},
vast number of experimental results and simulations are available for the three-dimensional case.
It was found that the for rectangular and/or square shaped capillaries there is a transition for
certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
The number is indicated in a number of works of as $\hat{Ca}=0.04$ \cite{cerro-bubble-train},
$\hat{Ca}=0.1$
\cite{cerro-space}, $\hat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger than
the critical capillary number, i.e. $Ca>\hat{Ca}$, then the flow becomes axisimmetric with the
radius of the droplet dependant on the capillary number.

The following pictures summarize the radius of deposited liquid on the capillary length.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
square capillaries are shown in Fig.\ref{fig:heil:three:dim}. On another hand authors performed
numerical simulations for rectangular shaped capillaries. It was found that aspect ratio $\alpha$
allows to put all the different curves on one purely empirical curve through the following parameter
$s_{\infty}=r_{\infty} \alpha^{-1/2}$, where $r_{\infty}$ is the aximsymmetric bubble radius
infinitely far behind the bubble tip and can be calculated from the flow rate in the computational
domain:
\begin{equation}
r_{\infty}=2\sqrt{(a-Q)/\pi}.
\end{equation}
It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
axisymmetric case. The $s_{\infty}$ dependency on $\alpha$ is shown in Fig.
\ref{fig:heil:three:dim}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/capillary_width_heil.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/capillary_rectangular_width_heil.eps}\\
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$.\label{fig:heil:three:dim}}
\end{figure}
There is also interesting thing to track down is the recirculating regime for flow fields. It is
known that for certain capillary number the streamlines either go around the bubble or create the
circulating regime in front of the bubble. The calculation ({\color{red} or analytical data? Look
at White 1991, pp. 119-120. As well put figures if there is a space from Heil et al.}) are presented
in Table \ref{table:recirculation:data}.
\begin{table}
\begin{tabular}{c|c}
$\alpha$& $Ca_{T}$\\
\hline
1& 0.691\\
1.1 & 0.688\\
1.3 & 0.666\\
1.5 & 0.631\\
\end{tabular}
\caption{The results for the recirculation region.\label{table:recirculation:data}}
\end{table}

\citet{shikazono-square} obtained a lot of different results for the deposition length versus the
capillary number variation for FC-40/air, ethanol/air and water/air experiments. All of them are
presented in Fig. \ref{fig:deposition:han}.
\begin{figure}
\includegraphics[width=0.33\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
\includegraphics[width=0.33\textwidth]{Figures/han_capillary_water_air.eps}\hfill
\includegraphics[width=0.33\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
\caption{Han results. \label{fig:deposition:han}}
\end{figure}

@@

\citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
shaped capillaries as square and triangle shaped capillaries. Some interesting results for bubble
shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
\ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
different capillary numbers. \ref{fig:wang:critical:capillary}.
\begin{figure}
\includegraphics[width=\textwidth]{Figures/wang_bubble_shape.eps}
\caption{The bubble shape across the length for different capillary numbers.
\label{fig:wang:bubble:shape}}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/wang_relative_velocity.eps}
\caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
velocity depending on the capillary number  \label{fig:wang:critical:capillary}}
\end{figure}

\section{Variation bubble length}
The work of \cite{heil-threedim} shows the variation of the bubble thickness through the length of
bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten}
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}
We want to obtain the same characteristics as indicated in Fig.
\ref{fig:thickness:variation:ca:ten}. One can see that the difference between diagonal radius and
usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
simulation is conducted by the grid $50\mathrm{x}50$ in the cross section that means that we will
see the difference in one grid node. Therefore the grid is $52\mathrm{x}52\mathrm{750}$. Here we
conduct a scaling:
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}
\end{equation}
The parameters for the binary liquid problem are chosen as $A=0.04$ and $k=0.01$. Therefore the
surface tension is:
\begin{equation}
\sigma = \sqrt{\frac{8 k A}{9}}=0.0596
\end{equation}
The relaxation times for liquid and gas are taken as, $\tau_{gas}=0.7$ and $\tau_{liq}=2.5$,
therefore the viscosities ratio is $10$.
\begin{equation}
\begin{aligned}
&10=\frac{\frac{2}{3} U_{bubble}}{0.0596},
&U_{bubble}=0.894
\end{aligned}
\end{equation}
which is quite high for the lattice Boltzmann method. We can decrease the ratio by scaling
parameters $k$ and $A$ while keeping the interface thickness. Let us choose parameters as $k=0.001$
and $A=0.004$, therefore the mobility should be adjusted to address the issue with small parameters
and small surface tension in this case. It parameters are chosen as $A=0.004$ and $k=0.001$ then
the velocity of the bubble is  $0.0894$. Therefore we can adjust the pressure gradient to plug it
into the bubble initialization:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca=0.0001907,
\end{equation}
which is quite high for lattice Boltzmann simulations. Therefore probably one more time the surface
tension needs to be reduced. {\color{red} Many authors state that increasing mobility kills the
proper results of simulations}. 
The film thickness as it is seen from the picture is taken as $70$ percent of the half channel.
Therefore the initalization width is $0.15*50=7.5$.

\section{Non-axisymmetric case}
As it was indicated before for a quite low capillary number the flow becomes axisymmetric. For
example, according to \cite{heil-threedim} for the capillary
number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
 resolved around $40-50$ percent for calculations to be convergent and grid independent. This
condition certainly implies a large grid with the computational high demand. For instance, to
properly simulate the abovementioned example the grid should be as 
$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
interface whether it can produce the consistent film thicknesses for low capillary numbers. The
interface occupies around $1-2$ lattice units.

@@

One can see the phase profile in the middle of the bubble in
Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
the calculation of the bubble velocity in the center of the bubble. The radiuses are as follows:
$R_{axis}=0.99$ and $R_{diag}=1.400$. While th  
1.4000714267

\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/}
\end{figure}

@@

\section{Steady-state case}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ with the initial
width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=3
\mathrm{x}10^{-5}$. The simulation was performed for $200000$,$400000$,$600000$ and $800000$ time
iterations. The results are summarised in Table \ref{table:steady:state}. One of the profiles in
the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}.
\begin{table}
\begin{tabular}{|c|c|c|c|}
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
$200000$&$0.285$&$0.873$&$0.921$\\
$400000$&$0.274$&$0.873$&$0.920$\\
$600000$&$0.263$&$0.872$&$0.919$\\
$800000$&$0.252$&$0.873$&$0.917$\\
\end{tabular}
\caption{Results for the steady-state case. One can see that $200000$ steps are enough to structure
the bubble. However, the interface is underresolved. That could be the influence of the bubble
length.\label{table:steady:state}}
\end{table}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
\caption{Slice in the middle of the bubble and
three-dimensional example. \label{fig:steady:state:profile:example}}
\end{figure}

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Recently we investigated the applicability and parameters choice for the two-dimensional case
Bretherton problem using the lattice Boltzmann method (LBM) \cite{kuzmin-binary2d}. This
paper is the continuation of the previous work and is focused on the three-dimensional case to
validate the LBM for simulation of phenomena for rectangular and square capillaries in the
moderate range of Capillary number parameters. 
\end{abstract}

\section{What to examine}
My plan is to investigate preferably the low capillary number regims as far we have interesting
phenomena occuring where: the axisymmetric radiuses for axes and diagonals. As far as it's
computationally very expensive to resolve the film thickness, it will be nice to investigate
whether underresolved films simulations can produce consistent results. The second thing to study
is the velocity pattern which is much easier to examine because the threshold is located for
$Ca\approx 0.6$. The third thing to study is the film thickness dependancy for the range of
capillary numbers. The last thing to study is the applicability of derived by \citet{heil-threedim}
correlations for rectangular channels of different shapes.

\section{Introduction}
While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-bretherton},
vast number of experimental results and simulations are available for the three-dimensional case.
It was found that the for rectangular and/or square shaped capillaries there is a transition for
certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection, see the
example in Fig. \ref{fig:quarter:capillary:capillary01} for a non-axisymmetric case.

The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the film
thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} performed three-dimensional simulations for circular shaped and for the
square shaped capillaries and gravitational effects for circular shaped capillaries. The results for
square capillaries are shown in Fig.\ref{fig:heil:three:dim}. Also, authors performed
numerical simulations for rectangular shaped capillaries. It was found that for different shaped
capillaries there is an empirical correlation allowing to put all the different curves for film
thickness dependency on the capillary number on one curve through introduction of the following
parameter
$s_{\infty}=r_{\infty} \alpha^{-1/2}$. The parameter $r_{\infty}$ is the aximsymmetric bubble radius
infinitely far behind the bubble tip and can be calculated as:
\begin{equation}
r_{\infty}=2\sqrt{(\alpha-Q)/\pi},
\end{equation}
where $Q$ is the net flow rate in the computational domain. {\color{red} What is it? - how to
calculate? Why velocity is involved?}

It is also indicated that for $\alpha\geq\hat{\alpha}=2.04$ the interface can not make up the
axisymmetric case for the range of capillary numbers. The $s_{\infty}$ dependency on $\alpha$ is
shown in Fig.
\ref{fig:heil:three:dim}.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/capillary_width_heil.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/capillary_rectangular_width_heil.eps}\\
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}

\citet{heil-threedim} indicate the interesting occuring thing in square and rectangular channels
as the recirculating regime for flow fields.  It is
known that for certain capillary number the streamlines either go around the bubble or create the
circulating regime in front of the bubble, Fig. \ref{fig:heil:recirculating}.  
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/recirculating.eps}
\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
Streamlines in the moving frame of reference for a rectangular tube $\alpha=1.5$ (a) Ca = 0.2,
(b) Ca = 0.5, (c) Ca = 0.55, (d) Ca = 2.0. Streamlines are shown on the surface of the bubble and
in the plane along the shorter semi-axis.
\label{fig:heil:recirculating}}
\end{figure}
The critical capillary numbers for changing velocity pattern depending on the eccentricity
parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
\begin{table}
\begin{tabular}{c|c}
$\alpha$& $Ca_{T}$\\
\hline
1& 0.691\\
1.1 & 0.688\\
1.3 & 0.666\\
1.5 & 0.631\\
\end{tabular}
\caption{The results for the recirculation region.\label{table:recirculation:data}}
\end{table}

\citet{shikazono-square} obtained a lot of different results for the deposition length versus the
capillary number variation ethanol/air and water/air experiments. All of them are
presented in Fig. \ref{fig:deposition:han}.
\begin{figure}
%\includegraphics[width=0.47\textwidth]{Figures/han_capillary_fc40_air.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/han_capillary_water_air.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/han_capillary_ethanol_air.eps}\\
\caption{The axial and diagonal radiuses dependency on the capillary number. Courtesy of
\citet{shikazono-square} \label{fig:deposition:han}}
\end{figure}

@@

\citet{wang-non-circular} performed numerical simulations by VOF technique to for the non-circular
shaped capillaries as square and triangle shaped capillaries. The VOF technique is the continuous
interface method based on the non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation. Some interesting results are
obtained for a bubble
shape depending on the capillary number, Fig. \ref{fig:wang:bubble:shape}. They also included
numerical simulations for the determination of the critical capillary number $\hat{Ca}$, Fig.
\ref{fig:wang:critical:capillary}. They also found the relative velocity for the bubble for
different capillary numbers.
\begin{figure}
\includegraphics[width=\textwidth]{Figures/wang_bubble_shape.eps}
\caption{The bubble shape across the length for different capillary numbers.
\label{fig:wang:bubble:shape}}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/wang_critical_capillary.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/wang_relative_velocity.eps}
\caption{Bubble radiuses variations over the range of capillary numbers (left) and the relative
velocity depending on the capillary number.  \label{fig:wang:critical:capillary}}
\end{figure}

\section{Non-axisymmetric case}
We first start with the non-axisymmetric case. As it was indicated before for a quite low capillary
number the flow becomes axisymmetric. For
example, according to \cite{heil-threedim} for the capillary
number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
 resolved around $40-50$ percent for calculations to be convergent and grid independent. This
condition certainly implies a large grid with the computational high demand. For instance, to
properly simulate the abovementioned example the grid should be as 
$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
interface whether it can produce the consistent film thicknesses for low capillary numbers. The
interface occupies around $1-2$ lattice units.

@@

One can see the phase profile in the middle of the bubble in
Fig.\ref{fig:quarter:capillary:capillary01}. The associated capillary number is $0.053$ based on
the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
are as $R_{axis}=0.93$ and $R_{diag}=1.06$. While $R_{axis}=0.93$ is underestimated parameter and
we can't take into the consideration apriori, one can consider parameter $R_{diag}=1.06$ the
parameters for the capillary number $0.053$ should be according to \citet{heil-threedim} as $0.97$
and $1.05$ according to the VOF simulations of \citet{wang-non-circular}. One can conclude that
even though the axis film thickness is not properly resolved the diagonal values are consistent and
can give an approximate information. 
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/quarter.eps}\\ %\hfill
\includegraphics[width=0.97\textwidth]{Figures/stream_cap01.eps}\\
\caption{Streamlines and phase profile of the quarter of the channel. The capillary number is
$0.053$ and the film thickness is underresolved. \label{fig:quarter:capillary:capillary01}}
\end{figure}
{\color{red} Add more data here either with better grid or with the less velocity that we can
track the numbers.}

\section{Variation bubble length}
The works of \citet{heil-threedim} of \citet{wang-non-circular} show the variation of the bubble
thickness through the length of
bubble. The length variation is presented in Fig. \ref{fig:thickness:variation:ca:ten} and
Fig.\ref{fig:wang:bubble:shape}. One can see that it takes a certain distance for the diagonal
radius to be equal to the axial even in the symmetric regime.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}
We want to obtain the same characteristics as indicated in Fig.
\ref{fig:thickness:variation:ca:ten}. One can see that the difference between diagonal radius and
usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
simulation is conducted by the grid $50\mathrm{x}50$ in the cross section that means that we will
see the difference in one grid node. Therefore the grid is $52\mathrm{x}52\mathrm{750}$. Here we
conduct a scaling:
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}
\end{equation}
The parameters for the binary liquid problem are chosen as $A=0.04$ and $k=0.01$. Therefore the
surface tension is:
\begin{equation}
\sigma = \sqrt{\frac{8 k A}{9}}=0.0596
\end{equation}
The relaxation times for liquid and gas are taken as, $\tau_{gas}=0.7$ and $\tau_{liq}=2.5$,
therefore the viscosities ratio is $10$.
\begin{equation}
\begin{aligned}
&10=\frac{\frac{2}{3} U_{bubble}}{0.0596},
&U_{bubble}=0.894
\end{aligned}
\end{equation}
which is quite high for the lattice Boltzmann method. We can decrease the ratio by scaling
parameters $k$ and $A$ while keeping the interface thickness. Let us choose parameters as $k=0.001$
and $A=0.004$, therefore the mobility should be adjusted to address the issue with small parameters
and small surface tension in this case. It parameters are chosen as $A=0.004$ and $k=0.001$ then
the velocity of the bubble is  $0.0894$. Therefore we can adjust the pressure gradient to plug it
into the bubble initialization:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2} \gamma Ca=0.0001907,
\end{equation}
which is quite high for lattice Boltzmann simulations. Therefore probably one more time the surface
tension needs to be reduced. {\color{red} Many authors state that increasing mobility kills the
proper results of simulations}. 
The film thickness as it is seen from the picture is taken as $70$ percent of the half channel.
Therefore the initalization width is $0.15*50=7.5$.

@@

\section{Steady-state case}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with $\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$,$150000$,$160000$, $170000$ and
$180000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate.

One of the
profiles in
the middle of the bubble is shown in Fig. \ref{fig:steady:state:profile:example}.
\begin{table}
\begin{tabular}{|c|c|c|c|}
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
$140000$&$0.0532$&$0.9379$&$1.075$\\
$150000$&$0.0533$&$0.9375$&$1.072$\\
$160000$&$0.0535$&$0.9372$&$1.069$\\
$170000$&$0.0536$&$0.9369$&$1.066$\\
$180000$&$0.0537$&$0.9366$&$1.062$\\
\end{tabular}
\caption{Results for the steady-state case. One can see that $200000$ steps are enough to structure
the bubble. However, the interface is underresolved. That could be the influence of the bubble
length.\label{table:steady:state}}
\end{table}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/bullet.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/example_crossection.eps}\\
\caption{Slice in the middle of the bubble and
three-dimensional example. \label{fig:steady:state:profile:example}}
\end{figure}
